---
title: "Assignment 9"
author: "Ev Berger-Wolf"
date: "2023-01-30"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

1. Using a `for` loop and an `if` statement, write a function to calculate the number of zeroes in a numeric vector. 
  - Before entering the loop, set up a vector of integers with some zero values. 
  - Then, set up a counter variable `counter <- 0`. 
  - Inside the loop, add 1 to counter each time you have a zero in the vector. 
  - Finally, use `return(counter)` for the output.
  
```{r}
my_vec <- sample(rep(c(0:5),each=5),size=10)


zeroCount <- function(in_vec=NULL){
  counter <- 0
  for(i in 1:length(in_vec)){
    if(in_vec[i] == 0) {counter <- counter + 1}
  }
  return(counter)
}

cat("vector =", my_vec, "\nnumber of zeros =", zeroCount(my_vec), "\n")
```

2. Write a function that takes as input two integers representing the number of rows and columns in a matrix.
  - The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
prodMat <- function(num_rows, num_cols) {
  mat_out <- matrix(nrow=num_rows, ncol=num_cols)
  for(i in 1:num_rows){
    for(j in 1:num_cols){
      mat_out[i,j] <- i*j
    }
  }
 return(mat_out) 
}
prodMat(sample(1:10,1),sample(1:10,1))
```

3. Use a `for` loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times.
  - Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows).




