---
title: "Assignment 5"
author: "Ev Berger-Wolf"
date: "2023-01-19"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#### Problem 1:

##### Create a list with the following named elements:

  - my_matrix, which is a 4 x 4 matrix filled with random uniform values
  - my_logical, which is a 100-element vector of TRUE or FALSE values. You can do this efficiently by setting up a vector of random values and then applying an inequality to it.
  - my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}
my_matrix <- matrix(runif(16),ncol=4,nrow=4)
my_logical <- runif(100)>0.5
# or my_logical <- as.logical(sample(0:1, 100, replace=T))
my_letters <- sample(letters)

my_list <- list(my_matrix, my_logical, my_letters) 
print(my_list)
```

##### Then, complete the following steps:

Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
my_list2 <- list(my_matrix[2,2],my_logical[2],my_letters[2])
print(my_list2)
```

Use the typeof() function to confirm the underlying data types of each component in this list

```{r}
for (i in 1:length(my_list2)) {
  print(typeof(my_list2[[i]]))
}
# or str(my_list2)
```

Combine the underlying elements from the new list into a single atomic vector with the c() function.

```{r}
my_vec <- c(my_list2[[1]],my_list2[[2]],my_list2[[3]])
print(my_vec)
```

What is the data type of this vector?

```{r}
# character
typeof(my_vec) 
```

#### Problem 2:

Create an empty data frame with two variables (= columns) and 26 observations (= rows) below

```{r}
d_frame <- data.frame(matrix(nrow = 26,ncol = 2))
head(d_frame)
```

Call the first column/variable my_unifs and fill it with 26 random uniform values from 0 to 10

```{r}
colnames(d_frame) <- c("my_unifs","")
d_frame["my_unifs"] <- runif(n=26,min=0,max=10)
head(d_frame)
```

Call the second variable my_letters and fill it with 26 capital letters in random order.

```{r}
colnames(d_frame) <- c(colnames(d_frame)[1],"my_letters")
d_frame["my_letters"] <- sample(LETTERS)
head(d_frame)
```

For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA

```{r}
d_frame[sample(1:dim(d_frame)[1],4),"my_unifs"] <- NA
sum(is.na(d_frame["my_unifs"]))
```

For the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
which(is.na(d_frame["my_unifs"]))
# or which(!complete.cases(d_frame))
```

Re-order the entire data frame to arrange the second variable in alphabetical order.

```{r}
d_frame <- d_frame[order(d_frame$my_letters),] 
#if you want opposite, decreasing = TRUE
head(d_frame)
```

Calculate the column mean for the first variable.

```{r}
mean(d_frame$my_unifs) #returns NA
mean(na.omit(d_frame$my_unifs)) #returns mean without NA
# or mean(d_frame$my_unifs, na.rm = T)
```

Rename the second column ‘my_LETTERS’

```{r}
colnames(d_frame) <- c(colnames(d_frame)[1],"my_LETTERS")
# or names(d_frame)[2] <- "my_LETTERS"
colnames(d_frame)
```

