---
title: "Assignment 8"
author: "Ev Berger-Wolf"
date: "2023-01-26"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#### Code Reflection

- Go through your recent assignments and the posted solutions - compare your code to the posted code.
- Are there differences between the posted code and yours? Did you notice you were missing anything?
Make sure you annotate your code using # and leave brief descriptions of your thought process.

(See Assignment 6 for updated annotations)

##### Weekly Assignment 2:

1. Data frame function

```{r}
### My solution:
dFrameMaker <- function(means=c(10,45),stdevs=c(1.5,2)) {
  d_frame <- data.frame(matrix(nrow = 50, ncol = 2)) #create data frame with 50 rows and 2 columns
  colnames(d_frame) <- c("Type","Value") #name the columns "Type" and "Value"
  
  d_frame["Type"] <- rep(c("Control","Treatment"),each=25) #Type: 25 Control, 25 Treatment
  d_frame["Value"] <- c(rnorm(n=25,mean=means[1],sd=stdevs[1]),
                        rnorm(25,mean=means[2],sd=stdevs[2])) #Value: 2 types of normal distributions
  
  return(d_frame) # returns the data frame
}


### Suggested solution:
simData <- function(mean1=10, sd1=1.5, mean2=45, sd2=2, n=25){
  
  Type <- rep(c("Control", "Treatment"), each = n)
  Value <- c(rnorm(n, mean = mean1, sd = sd1), rnorm(n, mean = mean2, sd = sd2))
  dframe <- data.frame(Type, Value)
  return(dframe)
}

```

Notes:

- I should create data frame after I create each column as vectors to improve efficiency
- I could improve flexibility by adding n as an argument
- Setting a seed ensures the same output each time

2. P-value function

```{r}
## My solution:

anovaTest2 <- function(d_frame=NULL){
  # Default data frame randomly generated from data frame maker
  if (is.null(d_frame)){
    d_frame <- dFrameMaker()
  }
  
  # if the data frame doesn't have 2 columns, return an error message
  if (length(d_frame)!=2){
    return("Function error: The data frame must have 2 columns.")
  }
  
  colnames(d_frame) <- c("Explanatory","Response") #rename column names
  my_anova <- aov(Response ~ Explanatory, data = d_frame) #run anova
  anova_tbl <- summary(my_anova) #extract anova summary
  pOut <- c(pVal=anova_tbl[[1]]["Explanatory","Pr(>F)"]) #extract p-value
  
  return(pOut) #output p-value
}

## Suggested solution:

myPval <- function(df=NULL){
  
  test <- aov(Value ~ Type, data=df)
  testSummary <- summary(test)
  p_value <- testSummary[[1]][["Pr(>F)"]][1]
  return(p_value)
}
```

Notes:

- The additions I made are not necessary, but maybe useful.
- The error message I implemented prevents unclear error messages generated by R

3. Testing my functions
  - Can you run your simulated data function with new values for the means? If not, what do you need to change about your code?

```{r}
data1 <- dFrameMaker(means=c(20,40), stdevs=c(3,5.5)) #using my function
data2 <- simData(mean1=20, mean2=40, sd1=3, sd2=5.5) #given function
data1
data2
```

```{r}
# they create the same p-values
anovaTest2(data1)
myPval(data1)

anovaTest2(data2)
myPval(data2)
```

My functions work!!



