---
title: "Assignment 7"
author: "Ev Berger-Wolf"
date: "2023-01-24"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r}
library(tidyverse) #includes dplyr & ggplot
library(ggthemes)
library(patchwork)
library(viridis)
```


#### Problem 1:
Using the built-in starwars data set in the dplyr package, create a new tibble with the following modifications:

  - It only includes the species that have more than 1 individual represented in the data set.
  - It only includes the columns that are not of type list

```{r}
# grab only species with more than one observation
data(starwars)
spec_counts <- starwars %>%
  group_by(species) %>%
  summarize(specCounts = n())
spec_names <- unlist(na.omit(spec_counts[spec_counts$specCounts > 1,"species"]))
spec_names
```

```{r}
# grab all column names that are not lists
col_vec <- c()
x<-1
for(i in 1:ncol(starwars)){
  if(typeof(starwars[[i]]) != "list"){
    col_vec[x] <- colnames(starwars)[i]
    x <- x+1
  }
}
col_vec
```

```{r}
# pipe in the parameters
starwars2 <- starwars %>%
  filter(species %in% spec_names) %>%
  select(all_of(col_vec)) %>%
  glimpse()
```

#### Problem 2:
Clean your data set: only include observations/rows that do not have an NA in the height column. Then, use dplyr to print a summary table showing the mean, median, and standard deviation of height for each species.

```{r}
# remove all NAs in height variable
starwars_clean <- starwars2 %>%
  filter(complete.cases(height)) %>%
  glimpse()
```

```{r}
# summarize mean, median, and standard deviation
summary_tbl <- starwars_clean %>%
  group_by(species) %>%
  summarize(meanHeight = mean(height), medianHeight = median(height), stdevHeight = sd(height)) %>%
  print()
```

#### Problem 3:

Create a box plot figure that shows the variation in heights among Star Wars species. 

  - Do the values in your summary table line up with the plot?
  - Code so that each box plot (Species) is be filled with a different color. 
  - Change your axes labels so that species is capitalized and height is capitalized and shows units (cm). 
  - Use the theme_minimal option. 
  - Then look into the annotate or geom_text functions and add the text “p=1.2E-0.5” to the top right corner.
  - Overlay the data points onto the box plots. See if you can change the color of the points, depending on the gender or sex variables.



```{r}
# boxplot based on species and height
newton <- ggplot(starwars_clean, aes(x=species, y=height, fill=species)) +
  geom_boxplot() +
  xlab("Species") +
  ylab("Height (cm)") +
  theme_minimal() + 
  geom_text(aes(x=tail(spec_names, n=1),y=max(height),label="p=1.2E-0.5"), size=3) +
  labs(title="Starwars data: Heights (cm) by Species") + 
  geom_point(aes(x=species, y=height, color=gender)) +
  scale_color_manual(values=c("green","red")) +
  guides(fill="none") 
newton #like fig(ure) newton!
```



#### Problem 4:

Using the same or a different built-in data set, create at least one additional (and different kind of) plot. Experiment with different arguments, features, and colors.

```{r}
data(iris) #load in iris data
colormappy <- c("#D81B60","#2196F3","#66BB6A") #set colors for each species
newton2 <- ggplot(iris, aes(x=Sepal.Length,y=Petal.Length,fill=Species,color=Species)) +
  geom_point(shape=sample(1:25,1)) + #random shape 
  scale_fill_manual(values = colormappy) + #fill and outline should be the same
  scale_color_manual(values = colormappy) +
  theme_bw() +
  xlab("Sepal Length (cm)") + ylab("Petal Length (cm)") +
  labs(title="Iris data: Petal vs. Sepal Length (cm)")
newton2
```

#### Problem 5:

Finally, create a multi-panel figure with the graphs you created.

```{r}
newton/newton2 + plot_layout(heights = c(2, 1)) #stacked plots, 1st plot is larger than the 2nd
```


