---
title: "Assignment 3"
author: "Ev Berger-Wolf"
date: "2023-01-12"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#### Problem 1:
Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the variable z and print the value stored in each z.

```{r}
x <- 1.1
a <- 2.2
b <- 3.3
```

##### A: $x^{a^b}$

```{r}
z <- x^(a^b)
print(z)
```

##### B: $(x^a)^b$

```{r}
z <- (x^a)^b
print(z)
```

##### C: $3x^3 + 2x^2 + 1$

```{r}
z <- 3*(x^3) + 2*(x^2) + 1
print(z)
```

#### Problem 2:
Using the `rep` and `seq` functions, create the following vectors:

##### A: $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$

```{r}
c(seq(from=1,to=8),seq(from=7,to=1))
# or c(1:8,7:1)
```

##### B: $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$

```{r}
rep(x=1:5,times=c(1:5))
# or rep(1:5,1:5)
```

##### C: $(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)$

```{r}
rep(x=5:1,times=c(1:5))
# or rep(5:1,1:5)
```

#### Problem 3: 
Create a vector using the code: `queue <- c("sheep", "fox", "owl", "ant")`, where queue represents the animals that are lined up to enter an exclusive club, with the sheep at the front of the line. Using R expressions, update `queue` as:

```{r}
queue <- c("sheep","fox", "owl", "ant")
print(queue)
```

##### A: the serpent arrives and gets in the back of the line

```{r}
queue <- c(queue,"serpent")
# or queue[length(queue)+1] = "serpent"
print(queue)
```

##### B: the sheep enters the club (so disappears from the line)

```{r}
queue <- queue[-which(queue=="sheep")]
print(queue)
```

##### C: the donkey arrives and talks his way to the front of the line

```{r}
queue <- c("donkey",queue)
print(queue)
```

##### D: the serpent gets impatient and leaves

```{r}
queue <- queue[-which(queue=="serpent")]
# or queue <- queue[queue != "serpent"]
print(queue)
```

##### E: the aphid arrives and the ant invites him to cut in line (hint: check out the append function)

```{r}
queue <- append(queue,"aphid",which(queue=="ant")-1)
print(queue)
```

##### F: Finally, determine the position of the aphid in the line (hint: use the which function)

```{r}
which(queue=="aphid")
```

