---
title: "Assignment 1"
author: "Ev Berger-Wolf"
date: "2023-01-10"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes 1/9/23

### This is a script. You can save code here to run again

Quick way to run code: Command + Enter

For example...

```{r addition}
2+2
```
Output should be in the Console

#### Variables show up in your Environment

Assign variables using var <- value

```{r variable assignment}
a <- 2+2
```

#### Terminal: Your shell, where you can speak directly to your computer
- Common shell command: ls --> Lists your directories
- setwd() - sets working directory
- getwd() - gets working directory
- Shortcut for help(): ?function or ??general_search
- *Always name your homepage index.Rmd

#### Terminal git commands to know
- git status -> tells you how up to date you are with GitHub
- git add -A -> adds untracked changes to the stage
- git commit -am 'message' -> commits all changes on the stage with the message
- git push -> Pushes changes to GitHub cloud

***

## Notes 1/10/23

#### git and GitHub
- Method of version control, which allows you to keep (& restore) past versions of files
- Also allow programmers to collaborate
- git: Keeps track of all the files; local version control
- GitHub: Free remote website for public data/project storage

#### Best formats for version control:
- Plain text (.txt)
- .R scripts
- Markdown (.md/.Rmd)
- html

#### Not great formats for version control (esp. in GitHub):
- PDFs
- Videos/pictures
- Word docs

#### Markdown: text formatting language used in multiple languages
- Type of markup language which allows you to format plain text
- R has its own version (RMarkdown)

#### GitHub repository:
- We can 'clone' (download a copy) the repo onto our local computer
- This clone can be edited locally, without changing the global version on the web
- Working on a repository = making changes to files, such as:
  - Adding a new file
  - Altering existing files
  - Deleting existing files
- Commit (noun): Snapshot of the current file system/local repository
  - You can return to previous commits and bring old files/changes back
  - Use messages to yourself to know which version you're committing
- Commit (verb): Saving and recording your changes locally
- Branches: Cloning a certain commit
  - Lets you try different things to eventually (maybe) merge the offshoot branch with the main branch
- Push: Transferring changes from local repo to GitHub repo
- Pull: Brings down GitHub repo to your local repo to sync them (helps with remote work and collaboration)

#### R Project:
- Organization system to divde your work into multiple contexts
- Each project has its own working directory
- Great workspace
- Can be linked to GitHub to access history and source documents


### Some formatting tricks:
# Header 1
## Header 2
### Header 3
This is plain text. You can also *italicize* and **bold** words.

`## this is where I can put anything. It's called code fencing!`

You can link things like [this!](file:///Users/ev/Downloads/jack-hamilton-EP_OHkgn1JI-unsplash.jpg)