---
title: "Assignment 9"
author: "Ev Berger-Wolf"
date: "2023-01-30"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

1. Using a `for` loop and an `if` statement, write a function to calculate the number of zeroes in a numeric vector. 
  - Before entering the loop, set up a vector of integers with some zero values. 
  - Then, set up a counter variable `counter <- 0`. 
  - Inside the loop, add 1 to counter each time you have a zero in the vector. 
  - Finally, use `return(counter)` for the output.
  
```{r}
my_vec <- sample(rep(c(0:5),each=5),size=10)


zeroCount <- function(in_vec=NULL){
  counter <- 0
  for(i in 1:length(in_vec)){
    if(in_vec[i] == 0) {counter <- counter + 1}
  }
  return(counter)
}

cat("vector =", my_vec, "\nnumber of zeros =", zeroCount(my_vec), "\n")
```

2. Write a function that takes as input two integers representing the number of rows and columns in a matrix.
  - The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
prodMat <- function(num_rows, num_cols) {
  mat_out <- matrix(nrow=num_rows, ncol=num_cols)
  for(i in 1:num_rows){
    for(j in 1:num_cols){
      mat_out[i,j] <- i*j
    }
  }
 return(mat_out) 
}
prodMat(sample(1:10,1),sample(1:10,1))
```

3. Use a `for` loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times.
  - Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows).


```{r}

dFrameMaker <- function(means=c(10,45),stdevs=c(1.5,2), each_col=25) {
  d_frame <- data.frame(matrix(nrow = each_col*2, ncol = 2)) #create data frame with 50 rows and 2 columns
  colnames(d_frame) <- c("Type","Value") #name the columns "Type" and "Value"
  
  d_frame["Type"] <- rep(c("Control","Treatment"),each=each_col) #Type: 25 Control, 25 Treatment
  d_frame["Value"] <- c(rnorm(n=each_col,mean=means[1],sd=stdevs[1]),
                        rnorm(n=each_col,mean=means[2],sd=stdevs[2])) #Value: 2 types of normal distributions
  
  return(d_frame) # returns the data frame
}


anovaTest2 <- function(d_frame=NULL){
  # Default data frame randomly generated from data frame maker
  if (is.null(d_frame)){
    d_frame <- dFrameMaker()
  }
  
  # if the data frame doesn't have 2 columns, return an error message
  if (length(d_frame)!=2){
    return("Function error: The data frame must have 2 columns.")
  }
  
  colnames(d_frame) <- c("Explanatory","Response") #rename column names
  my_anova <- aov(Response ~ Explanatory, data = d_frame) #run anova
  anova_tbl <- summary(my_anova) #extract anova summary
  pOut <- c(pVal=anova_tbl[[1]]["Explanatory","Pr(>F)"]) #extract p-value
  
  return(pOut) #output p-value
}

#for-loop to find p-value 100 times
boots_dframe <- data.frame(rep_num = rep(NA,100), p_val = rep(NA,100))
for(i in 1:100){
  boots_dframe[i,"rep_num"] <- i
  boots_dframe[i,"p_val"] <- anovaTest2()
}

head(boots_dframe)

```


