---
title: "Assignment 4"
author: "Ev Berger-Wolf"
date: "2023-01-17"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

Assign to the variable n_dims a single random integer between 3 and 10.

```{r}
n_dims <- runif(1, min = 3, max = 10)
n_dims <- round(n_dims)
print(n_dims)
```

Create a vector of consecutive integers from 1 to n_dims$^2$.

```{r}
my_vec <- c(1:n_dims^2)
print(my_vec)
```

Use the sample function to randomly reshuffle these values.

```{r}
my_vec2 <- sample(my_vec)
``` 

Create a square matrix with these elements.

```{r}
sq_mat <- matrix(my_vec2, nrow = n_dims, ncol = n_dims)
```

Print out the matrix.

```{r}
print(sq_mat)
```

Find a function in r to transpose the matrix.

```{r}
sq_mat2 <- t(sq_mat)
```

Print it out again and note how it has changed.

```{r}
print(sq_mat2) #the rows and columns are flipped
```

Calculate the sum and the mean of the elements in the first row and then the last row.

```{r}
# First row
sum(sq_mat[1,])
mean(sq_mat[1,])

# Last row
sum(sq_mat[n_dims,])
mean(sq_mat[n_dims,])
```

Set your code up so you can re-run it to create a matrix of a different size by only changing the n_dims value.

