---
title: "My Neuron Analysis"
author: "Ev Berger-Wolf"
date: "2023-02-02"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

### Neuron analysis from example data
Using the `nat` package, I chose to analyze and plot some of the sample neuron data that was included in the package. This dataset contains information about Drosophila (fly) olfactory projection neurons, obtained from https://doi.org/10.1016/j.cell.2007.01.040.

```{r, setup, message=FALSE}
# Setting up libraries
library(tidyverse)
library(ggplot2)
library(nat)
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```

First, let's load in the data
```{r}
data("Cell07PNs")
dros_data <- Cell07PNs #rename it to something more informative
length(dros_data) #20 neurons
head(dros_data) #let's take a peek at the information included
```
Glomeruli are a way to section groups of neurons. Neurons of the same glomerular class typically have similar structure, function, and/or location. We can look at the different glomerular classes of the neurons in this dataset. 

```{r}
summary(dros_data[,"Glomerulus"]) #number of each neurons in each glomerular class
```

Let's say we want to see if there is a difference in the number of segments between neurons in different glomeruli. We can start by extracting the number of segments for each neuron to see if they do generally differ.

```{r, fig.dim=c(3,3)}
segment_data <- data.frame(neuron_ID = c(1:length(dros_data)), num_segs = rep(NA, length(dros_data)))
for(i in 1:length(dros_data)){
  segment_data[i,"num_segs"] <- dros_data[[i]]$NumSegs
}

hist(segment_data$num_segs, breaks=20)
```

It looks like we have a generally bimodal distribution of the number of segments. Let's see if it significantly differs between glomeruli.

```{r, fig.dim=c(3,3)}
segment_data$glomerulus <- dros_data[,"Glomerulus"]
ggplot(segment_data, aes(x=glomerulus, y=num_segs, fill=glomerulus)) +
  geom_boxplot() +
  theme(legend.position="none")
```

At first glance, DP1m neurons seem to have a significantly different number of segments than the other glomeruli. Let's double check this with an ANOVA.

```{r}
segment_data$glomerulus <- factor(segment_data$glomerulus)
glom_aov <- aov(num_segs ~ glomerulus, data=segment_data)
summary(glom_aov)
```

Now that we know that there is at least one signifcantly different glomerulus, let's conduct a post-hoc test to determine which of the glomeruli significantly differ in segment number.

```{r}
TukeyHSD(glom_aov)
```
Here, we can confirm that it is the DP1m glomerulus that has a significantly different number of segments. Let's take a look at the morphology of the DP1m neurons versus the rest.


```{r}
plot(dros_data, subset=Glomerulus!="DP1m", col='red', 
     WithNodes=F, main="DP1m vs. all other neurons") 
plot(dros_data, subset=Glomerulus=="DP1m", add=TRUE, col="blue", WithNodes=F)
legend('bottomright', legend = c("Not DP1m","DP1m"), fill=c("red","blue"))
```

**Now we can see the qualitative difference in the number of segments of DP1m neurons from non-DP1m neurons. And that's one example of what the `nat` package can be used for!**
