---
title: "nat: The Neuroanatomy Toolbox for R"
author: "Ev Berger-Wolf"
date: "2023-02-02"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## Final Project: The Neuroanatomy Toolbox (nat) Package

The `nat` package is the core of the "natverse", a series of packages crucial for neuroanatomical visualization in R. It is highly effective for examining the structure of connectivity of neuronal circuits in various species. 

```{r, setup, message=FALSE}
# Setting up libraries
library(nat)
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```


#### Introduction:

`nat` creates 2 new classes of objects (aside from previous classes, such as `numeric`, `character`, etc.):

- `neuron`: individual neuron objects (see below)
- `neuronlist`: list of neuron objects

```{r}

## Let's load some  example data:
data("Cell07PNs") #olfactory projection neurons from https://doi.org/10.1016/j.cell.2007.01.040
class(Cell07PNs) #neuronlist (nat) and list (base R)

```

Because neuronlists are also lists, we can apply some functions to them that we can to lists.

```{r}
length(Cell07PNs) #number of objects (neurons) in this dataset
```

We can also index them similarly.

```{r}
class(Cell07PNs[[1]]) #neuron objects are also lists
```


#### Neurons:

Neuron objects have numerous fields. The most important fields are the following:

  - `d`: a data frame object describing the x,y,z positions of each of the nodes (projections) of the neuron and the diameter (w) of the projection
  - `NumPoints`: number of points in general
  - `StartPoint`, `BranchPoints`, `EndPoints`: all of the neuron's nodes
  - `SegList`: Vertex indices of the neuron's segments

Let's pull out the first `neuron` in the `neuronlist`.

```{r}
neur1 <- Cell07PNs[[1]]

# We can look at a summary of the neuron using summary.neuron (masked as summary)
summary(neur1)
```


#### Neuronlists:

- Primarily store all of the neurons in a dataset/subset.
- Can have an optional attached `data.frame` containing additional information about each of the neurons in the list, called metadata
- We can extract the metadata using `as.data.frame()`

```{r}
meta_data <- as.data.frame(Cell07PNs)
head(meta_data) #gives us a sense of the qualities we can subset by
```

Using metadata to index:

```{r}
# Let's say we want to know all of the glomeruli represented in the dataset
with(Cell07PNs, table(Glomerulus)) #shows number of neurons in each glomerulus

# Now we can subet neurons based on their glomerulus
va1d_n <- subset(Cell07PNs, Glomerulus=="VA1d") #new sub-neuronlist of neurons in glomerulus VA1d

# If we want a summary of the meta data, rather than the neuronlist:
summary(Cell07PNs[,]) #[,] accesses the data frame without as.data.frame()
```


#### Plotting neurons:

We can plot an individual neuron:

```{r, fig.dim=c(3,3)}
plot(neur1) #purple node = root/soma of the neuron, red nodes = branch points, green nodes = end points
plot(neur1, WithText=T) #we can label nodes based on their PointNo index from the data.frame d (see above)
plot(neur1, col="blue") #we can change the color of a neuron's axon using col

```

We can also plot multiple neurons:

```{r, fig.dim=c(3,3)}
plot(Cell07PNs[1:4]) #plots first 4 neurons from the neuronlist
plot(Cell07PNs) #plots all of the neurons
```

We can use subsetting to highlight a specific quality of the neurons:

```{r, fig.dim=c(3,3)}
plot(Cell07PNs, subset=Glomerulus!="VA1d", col='grey', 
     WithNodes=F, main="VA1d neurons") #main=title, removes nodes, plots all but the VA1d neurons in grey
plot(Cell07PNs, subset=Glomerulus=="VA1d", add=TRUE) #adds the Va1d neurons to the plot (superimposes using add=TRUE)
```

#### 3D Plots:

3D plotting is super important in anatomical mapping, since our bodies (specifically our brains) do not exist on a 2D plane.

- `nat` uses the `rgl` package to plot 3D models of the neurons (remember that the neuron has z coordinates too!).  

You can drag the neuron below around and zoom in and out.

```{r, webgl=TRUE, fig.dim=c(3,3)}
clear3d()
plot3d(neur1, col='blue')
```

You can also plot multiple neurons on a 3D plot.
```{r, webgl=TRUE, fig.dim=c(3,3)}
plot3d(Cell07PNs, subset=grepl("^D", Glomerulus), #subset all neurons with glomeruli starting with "D"
            col=Glomerulus, lwd=2, WithNodes=FALSE) #color by glomerulus, lwd = line weight
```

Access the nat package [here](https://github.com/natverse/nat).  
Information obtained from [https://natverse.org/nat/articles/neurons-intro.html](https://natverse.org/nat/articles/neurons-intro.html).

