---
title: "Weekly Assignment 4"
author: "Ev Berger-Wolf"
date: "2023-01-31"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
```{r, message=FALSE}
# Load libraries
library(tidyverse)
library(ggplot2)
library(ggthemes)
```

Simulate a brand new data frame (the type of data is up to you) and conduct a statistical analysis in R using 2 variables. Create a basic ggplot that goes with the analysis.

```{r, warning=FALSE}
## Simulating data: 
# Is there a difference in the number of spikes of inhibitory and excitatory neurons during rest vs. a memory task?
spike_data <- data.frame(neuron.type = rep(c("excitatory","inhibitory"), each=25), 
                         task = rep(c("memory","rest"), times=25))
set.seed(459)
for(i in 1:dim(spike_data)[1]){
  if(spike_data[i,"neuron.type"] == "excitatory"){
    m <- 50; stdev <- 8} else {
      m <- 25; stdev <- 12}
  if(spike_data[i,"task"] == "memory"){
    m <- m * 1.25 } else {
      m <- m * 0.75 }
  spike_data[i,"spike.num"] <- abs(round(rnorm(1, mean=m, sd=stdev)))
}

spike_data$neuron.type <- factor(spike_data$neuron.type)
spike_data$task <- factor(spike_data$task)
spike_sum <- spike_data %>%
  group_by(neuron.type, task) %>%
  summarize(meanSpike = mean(spike.num), sdSpike = sd(spike.num)) %>%
  print()
```

```{r}
## 2-way ANOVA
spike_aov <- aov(spike.num ~ neuron.type * task, data = spike_data)
summary(spike_aov)
```
There seems to be a significant difference in spike number between excitatory and inhibitory neurons, as well as between the memory task and rest. However, there does not seem to be a significant interaction.

```{r}
## Post-hoc multiple comparisons test
TukeyHSD(spike_aov)
```
Let's graph this!

```{r}
ggplot(spike_sum,aes(x = task, y= meanSpike, fill=neuron.type)) +
  geom_bar(stat="identity", position="dodge") +
  geom_errorbar(aes(ymin=meanSpike-sdSpike,
                    ymax=meanSpike+sdSpike), position=position_dodge2(padding=0.8), linewidth=0.35) +
  scale_fill_manual(values=c("#42A5F5","#81C784"),name = "Neuron Type") +
  labs(x="Task",y="Mean Number of Spikes") +
  theme_bw()
```

Neurons in this simulated data have significantly more spikes during a memory task ($\Delta$M = `r abs(signif(mean(unlist(spike_sum[spike_sum$task=="memory","meanSpike"]))-mean(unlist(spike_sum[spike_sum$task=="rest","meanSpike"])), digits=4))`, $\Delta$SD = `r abs(signif(mean(unlist(spike_sum[spike_sum$task=="memory","sdSpike"]))-mean(unlist(spike_sum[spike_sum$task=="rest","sdSpike"])),digits=4))`, *p* < .01), as well as significantly more spikes if they are excitatory, rather than inhibitory ($\Delta$M = `r abs(signif(mean(unlist(spike_sum[spike_sum$neuron.type=="excitatory","meanSpike"]))-mean(unlist(spike_sum[spike_sum$neuron.type=="inhibitory","meanSpike"])), digits=4))`, $\Delta$SD = `r abs(signif(mean(unlist(spike_sum[spike_sum$neuron.type=="excitatory","sdSpike"]))-mean(unlist(spike_sum[spike_sum$neuron.type=="inhibitory","sdSpike"])),digits=4))`, *p* < .01).





