---
title: "Weekly Assignment 2"
author: "Ev Berger-Wolf"
date: "2023-01-19"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

1. Create a function that returns a data frame with 2 columns (named “Type” and “Value”) and 50 rows/observations.
  - The first column should have “Control” as the first 25 observations and “Treatment” as the second half of observations.
  - The second column should have the first 25 values as random and normally distributed with a mean of 10 and standard deviation of 1.5. The next 25 values of the second column should be random and normally distributed with a mean of 45 and standard deviation of 2.
  - You can do this all as a single line of code in the function or by breaking it up into multiple code blocks.

```{r}
# FUNCTION: dFrameMaker
# Create a data frame of simulated data
# inputs: vector of means (means) and a vector of standard deviations (stdevs) for control and treatment.
#   Default is means = [10, 45]; sd = [1.5, 2]
# returns: the data frame created 

dFrameMaker <- function(means=c(10,45),stdevs=c(1.5,2)) {
  d_frame <- data.frame(matrix(nrow = 50, ncol = 2)) #create data frame with 50 rows and 2 columns
  colnames(d_frame) <- c("Type","Value") #name the columns "Type" and "Value"
  
  d_frame["Type"] <- rep(c("Control","Treatment"),each=25) #Type: 25 Control, 25 Treatment
  d_frame["Value"] <- c(rnorm(n=25,mean=means[1],sd=stdevs[1]),
                        rnorm(25,mean=means[2],sd=stdevs[2])) #Value: 2 types of normal distributions
  
  return(d_frame) # returns the data frame
}

```


2. Save your new function’s output as a variable, and use a function to view the first 6 rows of the data frame.

```{r}

my_frame <- dFrameMaker()
head(my_frame)

```


3. Let’s say you would like to know whether there is a statistically significant difference in “Value” (response variable y) depending on “Type” (explanatory variable x). Type ?aov in the console to determine how to run an analysis of variance (ANOVA) on your simulated data. Write a line of code that displays a summary of your ANOVA.

```{r}
my_anova <- aov(Value ~ Type, data = my_frame)
summary(my_anova)
```

4. Finally create a function that uses a data frame as its input and returns only the p-value of an ANOVA summary table (feel free to use Google/Stack Overflow). Write your code in such a way that you can use any simulated data set with two columns as the function’s argument.

```{r}
# FUNCTION: anovaTest2
# Find the p-value of an ANOVA for any data frame with 2 columns
# inputs: Some data frame (d_frame) with 2 columns
#   The 1st column should be the explanatory value and the 2nd column should be the response variable
# outputs: p-value based on the ANOVA for the 2 variables

anovaTest2 <- function(d_frame=NULL){
  # Default data frame randomly generated from data frame maker
  if (is.null(d_frame)){
    d_frame <- dFrameMaker()
  }
  
  # if the data frame doesn't have 2 columns, return an error message
  if (length(d_frame)!=2){
    return("Function error: The data frame must have 2 columns.")
  }
  
  colnames(d_frame) <- c("Explanatory","Response") #rename column names
  my_anova <- aov(Response ~ Explanatory, data = d_frame) #run anova
  anova_tbl <- summary(my_anova) #extract anova summary
  pOut <- c(pVal=anova_tbl[[1]]["Explanatory","Pr(>F)"]) #extract p-value
  
  return(pOut) #output p-value
}

anovaTest2() #run default
new_frame <- dFrameMaker(means=c(20,25),stdevs=c(5,7)) #create new data frame
anovaTest2(new_frame) #run ANOVA on the new frame

```

